---
title: "calc ssGSEA"
author: "jmzeng1314@163.com"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message  = F) 
```

# Arrangement of rpkm matrix

```{r}
###Matrix Information Extraction
a <- read.table('GSE112996_merged_fpkm_table.txt.gz',
                header = T,
                row.names=1)
raw_data<- a[,-1]
###Phenotype Information Extraction
pheno <- read.csv(file = 'GSE112996_series_matrix.txt')
pheno <- data.frame(num1 = strsplit(as.character(pheno[42,]),split='\t')[[1]][-1],
                    num2 = gsub('patient: No.','P',strsplit(as.character(pheno[51,]),split='\t')[[1]][-1]))
table(pheno$num2)
####data filtering
data<- a[!apply(raw_data,1,sum)==0,] 
####Remove rows with duplicate gene names, normalize
uni_matrix<- data[!duplicated(data$GeneName),]
rownames(uni_matrix) <- uni_matrix[,1]
uni_matrix <- uni_matrix[,-1]
uni_matrix <- log2(uni_matrix+1)
colnames(uni_matrix)<- gsub('X','',gsub('\\.','\\-',colnames(uni_matrix)))
uni_matrix<- uni_matrix[,order(colnames(uni_matrix))]
uni_matrix[1:4,1:4]
```

The resulting uni_matrix is our expression matrix, which is a data frame in R.

# Organize 28 immune-related gene sets

Extract the immune-related gene set from the pdf or Excel table, and read it into R to become a list object, as follows;

```{r}
library(GSVA)
library(Biobase)
load('gene_set.Rdata')
str(list)
```

# Run gsva on expression matrices and gene sets
```{r}
gsva_matrix<- gsva(as.matrix(uni_matrix), list,method='ssgsea',kcdf='Gaussian',abs.ranking=TRUE, verbose=F)
# DT::datatable(gsva_matrix)
```

There are two parameters to note here, the first is the method can be selected：
- gsva (Hänzelmann et al, 2013) 
- ssgsea (Barbie et al, 2009)
- zscore (Lee et al, 2008)
- plage (Tomfohr et al, 2005)

Then kcdf can choose


# Heatmap to visualize gsva results

```{r}
library(pheatmap)
gsva_matrix1<- t(scale(t(gsva_matrix)))
gsva_matrix1[gsva_matrix1< -2] <- -2
gsva_matrix1[gsva_matrix1>2] <- 2
normalization<-function(x){
  return((x-min(x))/(max(x)-min(x)))}
nor_gsva_matrix1 <- normalization(gsva_matrix1)
annotation_col = data.frame(patient=pheno$num2)
rownames(annotation_col)<-colnames(uni_matrix)
bk = unique(c(seq(0,1, length=100))) 
pheatmap(nor_gsva_matrix1,show_colnames = F,
         cluster_rows = F,cluster_cols = F,annotation_col = annotation_col,
         breaks=bk)#cellwidth=5,cellheight=5,fontsize=5 
```

# Divide the immune gene set into two categories to see the correlation

```{r}
anti_tumor <- c('Activated CD4 T cell', 'Activated CD8 T cell', 'Central memory CD4 T cell	', 'Central memory CD8 T cell', 'Effector memeory CD4 T cell', ' Effector memeory CD8 T cell', 'Type 1 T helper cell', 'Type 17 T helper cell', 'Activated dendritic cell', 'CD56bright natural killer cell', 'Natural killer cell', 'Natural killer T cell')
pro_tumor <- c('Regulatory T cell', 'Type 2 T helper cell', 'CD56dim natural killer cell', 'Immature dendritic cell', 'Macrophage', 'MDSC', 'Neutrophil', 'Plasmacytoid dendritic cell')
load('score.Rdata')
anti<- as.data.frame(gsva_matrix1[gsub('^ ','',rownames(gsva_matrix1))%in%anti_tumor,])
pro<- as.data.frame(gsva_matrix1[gsub('^ ','',rownames(gsva_matrix1))%in%pro_tumor,])
anti_n<- apply(anti,2,sum)
pro_n<- apply(pro,2,sum)
patient <- pheno$num2[match(colnames(gsva_matrix1),pheno$num1)]
library(ggplot2)
data <- data.frame(anti=anti_n,pro=pro_n,patient=patient)
anti_pro<- cor.test(anti_n,pro_n,method='pearson')
gg<- ggplot(data,aes(x = anti, y = pro),color=patient) + 
  xlim(-15,12)+ylim(-15,10)+
  labs(x="Anti-tumor immunity", y="Pro-tumor suppression") +
  geom_point(aes(color=patient),size=3)+geom_smooth(method='lm')+
  annotate("text", x = -5, y =7.5,label=paste0('R=',round(anti_pro$estimate,4),'\n','p<0.001'))
print(gg)
# ggsave(gg,filename = 'cor.png')
```



